TensorAlloy: an automatic atomistic neural network program for alloys

### 2019-10-28 doc/papers/nn/manuscript.tex
Slightly improves the manuscript to answer the questions of the last reviewer.

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. The symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help.  
In this work, we propose a virtual-atom approach to model the total energy of 
symmetry function descriptors based atomistic neural network. Our approach 
creates the computation graph directly from atomic positions. Thus, the 
derivations of forces and virial can be handled by TensorFlow automatically and 
efficiently. This new approach is implemented in our open-source program 
TensorAlloy, which supports constructing machine learning interaction potentials 
for both \textit{molecules} and \textit{solids}. The QM7 and SNAP/Ni-Mo datasets 
are used to demostrate the performances of our program.
\end{abstract}

@@

In this work, we propose a new TensorFlow-dependent apporach (the virtual atom 
approach, VAP) to model the total energy expression of the symmetry function 
descriptor based neural network. 
Our apporach can move the entry of the computation graph from atomic descriptors 
back to the raw atomic positions. Thus, calculating the derivatives of total 
energy with respect to atomic positons (force) or lattice tensor (virial) can be 
handled by TensorFlow directly and automatically.

@@

\begin{table}[h]
\centering
\begin{tabular}{ccc|cc|c}
\hline
  & \multicolumn{2}{c|}{Energy (meV/atom)} & \multicolumn{2}{c|}{Force (eV/\AA)}
  & Stress (GPa) \\
\hline
Elementary Model & Mo & Ni & Mo & Ni & Mo \\
\hline
Mo & 4.5 (13.2) & & 0.19 (0.25) & & 0.28 (0.87) \\
Ni & & 1.3 (1.2) & & 0.04 (0.05) & \\
\hline
\end{tabular}
\caption{\label{table:MAE1}
Comparion of the MAEs in predicted energies (mev/atom), forces (eV/\AA) and 
stress (GPa) relative to DFT for the two elementary models (Ni, Mo) and their
corresponding SNAP benchmarks (bracket)}
\end{table}

Table \ref{table:MAE} and \ref{table:MAE1} summarizes the energy, force and 
stress prediction performances of our trained models compared with their 
corresponding SNAP benchmarks. 
The original SNAP formalism contains both radial and angular 
interactions. However, in most cases, the radial-only TensorAlloy-trained 
models can significantly outperform corresponding SNAP models. The trainings of 
these models are also very fast. All these models are trained within 2 GPU hours
on our workstation.

>>>>

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. The symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help.  
In this work, we propose a virtual-atom approach to model the total energy of 
symmetry function descriptors based atomistic neural network. Our approach 
creates the computation graph directly from atomic positions. Thus, the 
derivations of forces and virial can be handled by TensorFlow automatically and 
efficiently. The virtual atom approach with AutoGrad within TensorFlow allows 
for efficient training to not just energies and forces, but also virial stress. 
This new approach is implemented in our open-source program TensorAlloy, which 
supports constructing machine learning interaction potentials for both 
\textit{molecules} and \textit{solids}. The QM7 and SNAP/Ni-Mo datasets are used 
to demostrate the performances of our program.
\end{abstract}

@@

In this work, we propose a new TensorFlow-dependent apporach (the virtual atom 
approach, VAP) to model the total energy expression of the symmetry function 
descriptor based neural network. 
Our apporach can move the entry of the computation graph from atomic descriptors 
back to the raw atomic positions. Thus, calculating the derivatives of total 
energy with respect to atomic positons (force) or lattice tensor (virial) can be 
handled by TensorFlow directly and automatically. Hence, the technical barrier 
of training ANNs with atomic forces and (especially) virial stress becomes 
completely neglectable.

@@

\begin{table}[h]
\centering
\begin{tabular}{ccc|cc|c}
\hline
  & \multicolumn{2}{c|}{Energy (meV/atom)} & \multicolumn{2}{c|}{Force (eV/\AA)}
  & Stress (GPa) \\
\hline
Elementary Model & Mo         & Ni        & Mo            & Ni   & Mo          \\
\hline
Mo               & 4.5 (13.2) &           & 0.19 (0.25)   &      & 0.28 (0.87) \\
Ni               &            & 1.3 (1.2) & & 0.04 (0.05) & 2.05               \\
\hline
\end{tabular}
\caption{\label{table:MAE1}
Comparion of the MAEs in predicted energies (mev/atom), forces (eV/\AA) and 
stress (GPa) relative to DFT for the two elementary models (Ni, Mo) and their
corresponding SNAP benchmarks (bracket)}
\end{table}

Table \ref{table:MAE} and \ref{table:MAE1} summarizes the energy, force and 
stress prediction performances of our trained models compared with their 
corresponding SNAP benchmarks. 
The original SNAP formalism contains both radial and angular 
interactions. However, in most cases, the radial-only TensorAlloy-trained 
models can significantly outperform corresponding SNAP models. We can also 
notice that including virial stress in the total loss function can significantly
reduce the stress prediction error. The virial stress 
(Equation \ref{eq:stress}) consists of two parts: the net force contribution 
$-F^TR$ and the partial force contribution 
$\left(\partial E / \partial \mathbf{h}\right)^T\mathbf{h}$. The atomic forces 
used by the total loss equation \ref{eq:loss} are net forces only but virial 
stress only depends on the partial forces (the force between two atoms). 
Compared with the binary \textbf{NN}(Mo-Ni) model, the elementary 
\textbf{NN}(Ni) model has lower net force error (0.04 eV/\AA vs 0.07 eV\AA). 
However, the elementary model trained with energy and forces only has much 
larger stress MAE (2.05 GPa vs 0.83 GPa). These results clearly suggest that 
including virial stress in the total loss function is necessary for obtaining 
better virial stress accuracy.
